---
title: "Notes"
output: html_document
date: "2024-09-26"
---

```{r load-packages, message=FALSE}
library(tidyverse) 
```

## Types
double 1, 2.44\
logical TRUE, FALSE \
integer 6L, 2:6\
character "k", "hello" \

```{r}
typeof("look")
typeof(2)
typeof(4L)
```

function c() to produce vector of any type\
```{r}
x <- c(1,2,3)
x
typeof(x)
y <- c("hello", "world")
y
typeof(y)
```

express one type as another with as.character(), as.numeric(), as.integer(), as.logical() \
data type hierarchy: character>double>integer>logical \
NaN - not a number \
Inf and -Inf - positive and negative infinity \
NA - not available \


# Data classes
```{r}
M1 <- matrix(c(1, 4, 3, 6, 7, 2), nrow=2, ncol=3)
M1
M2 <- matrix(c(1, 4, 3, 6, 7, 2), nrow=3, ncol=2)
M2
M1 %*% M2 # matrix multiplication
```

```{r}
l <- list(
  item1 = 1:4,
  item2 = "hi",
  item3 = list(
    item31 = 3
  ),
  item4 = c(TRUE, FALSE),
  item5 = diag(3)
)
l
```
data frmae is a list where all items have same lenght \

```{r}
df <- data.frame(
  x = c(1,2,3),
  y = c(TRUE, FALSE, TRUE),
  z = c("z", "x", "c")
)
df
typeof(df)
class(df)
```

use factors to handle categorical variables \
```{r}
x <- factor(c("BS", "MS", "PhD", "BS"))
x
class(x)
glimpse(x)
as.integer(x)
```
lubridate to work with dates easier \